---
title: "Riddler Classic - Chocolates"
output: html_notebook
---

From mathematician (and author of Basic Probability: What Every Math Student Should Know) Henk Tijms comes a choice matter of chance and chocolate:

I have 10 chocolates in a bag: Two are milk chocolate, while the other eight are dark chocolate. One at a time, I randomly pull chocolates from the bag and eat them — that is, until I pick a chocolate of the other kind. When I get to the other type of chocolate, I put it back in the bag and start drawing again with the remaining chocolates. I keep going until I have eaten all 10 chocolates.

For example, if I first pull out a dark chocolate, I will eat it. (I’ll always eat the first chocolate I pull out.) If I pull out a second dark chocolate, I will eat that as well. If the third one is milk chocolate, I will not eat it (yet), and instead place it back in the bag. Then I will start again, eating the first chocolate I pull out.

What are the chances that the last chocolate I eat is milk chocolate?

```{r}
library(tidyverse)
library(purrr)
```


```{r}
#Initial Settings
milk <- 2
dark <- 8
milk_prob <- milk/(milk+dark)
dark_prob <- dark/(milk+dark)
start_again <- TRUE
previous_draw <- ""
last_item <- "NA"

chocolate_fun <- function(milk, dark, milk_prob, dark_prob, start_again, previous_draw, last_item) {
  while(last_item == "NA")
    {
      while(start_again)
        {
          #Dealing with the first draw
          first_draw <- sample(x = c("milk","dark"), size = 1, prob = c(milk_prob, dark_prob))
          milk <- case_when(first_draw == "milk" ~ milk - 1,
                            TRUE ~ milk - 0)
          dark <- case_when(first_draw == "dark" ~ dark - 1,
                            TRUE ~ dark - 0)
          milk_prob <- milk/(milk+dark)
          dark_prob <- dark/(milk+dark)
          previous_draw <- first_draw
          start_again <- FALSE
        }
        #Dealing with subsequent draws
        next_draw <- sample(x = c("milk","dark"), size = 1, prob = c(milk_prob, dark_prob))
        milk <- case_when(next_draw == previous_draw & previous_draw == "milk" ~ milk - 1, TRUE ~ milk + 0)
        dark <- case_when(next_draw == previous_draw & previous_draw == "dark" ~ dark - 1, TRUE ~ dark + 0)
        milk_prob <- milk/(milk+dark)
        dark_prob <- dark/(milk+dark)
        start_again <- next_draw != first_draw
        previous_draw <- case_when(start_again == TRUE ~ "", TRUE ~ next_draw)
        last_item <- case_when(dark + milk == 1 & dark > 0 ~ "dark", 
                               dark + milk == 1 & milk > 0 ~ "milk",
                               dark == 0 ~ "milk",
                               milk == 0 ~ "dark",
                               TRUE ~ "NA")
  }
  return(last_item)
}

set.seed(1975)
outcomes <- replicate(1000, chocolate_fun(milk, dark, milk_prob, dark_prob, start_again, previous_draw, last_item))
```


```{r}
#summarize outcomes from trial 2

outcomes %>% 
  as_tibble() %>% 
  mutate(milk_outcome = case_when(value == "milk" ~ 1, TRUE ~ 0)) %>% 
  summarize(milk_outcome = sum(milk_outcome)/n())

outcomes %>% 
  as_tibble() %>% 
  group_by(value) %>% 
  count()
```


```{r}
#same concept as above without a function and using a for loop to build a vector of outcomes
for (i in 1:1000) {
  
milk <- 2
dark <- 8
milk_prob <- milk/(milk+dark)
dark_prob <- dark/(milk+dark)
start_again <- TRUE
previous_draw <- ""
last_item <- "NA"

  while(last_item == "NA")
    {
      while(start_again)
        {
          #Dealing with the first draw
          first_draw <- sample(x = c("milk","dark"), size = 1, prob = c(milk_prob, dark_prob))
          milk <- case_when(first_draw == "milk" ~ milk - 1,
                            TRUE ~ milk - 0)
          dark <- case_when(first_draw == "dark" ~ dark - 1,
                            TRUE ~ dark - 0)
          milk_prob <- milk/(milk+dark)
          dark_prob <- dark/(milk+dark)
          previous_draw <- first_draw
          start_again <- FALSE
        }
        #Dealing with subsequent draws
        next_draw <- sample(x = c("milk","dark"), size = 1, prob = c(milk_prob, dark_prob))
        milk <- case_when(next_draw == previous_draw & previous_draw == "milk" ~ milk - 1, TRUE ~ milk + 0)
        dark <- case_when(next_draw == previous_draw & previous_draw == "dark" ~ dark - 1, TRUE ~ dark + 0)
        milk_prob <- milk/(milk+dark)
        dark_prob <- dark/(milk+dark)
        start_again <- next_draw != first_draw
        previous_draw <- case_when(start_again == TRUE ~ "", TRUE ~ next_draw)
        last_item <- case_when(dark + milk == 1 & dark > 0 ~ "dark", 
                               dark + milk == 1 & milk > 0 ~ "milk",
                               dark == 0 ~ "milk",
                               milk == 0 ~ "dark",
                               TRUE ~ "NA")
  }
    ov <- append(last_item, ov)
    i <- i + 1
}
```


```{r}
#summarize outcomes from trial 2
ov %>% 
  as_tibble() %>% 
  mutate(milk_outcome = case_when(value == "milk" ~ 1, TRUE ~ 0)) %>% 
  summarize(milk_outcome = sum(milk_outcome)/n())

ov %>% 
  as_tibble() %>% 
  group_by(value) %>% 
  count()
```




